Homicide Increases in the COVID-19 Era

### 2025-03-04 2_SecondDraft/Draft2.qmd
Submitted draft version

author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
fontsize: 10pt
toc: true                   # add table of contents at the beginning
toc-depth: 3               # Only titles that start with # or ##
geometry: margin=1in
bibliography: references.bib
format:
  pdf: default
number-sections: true
---

@@

# primary datasets
# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")

crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")

# primary dataset zip code and geometry data
crimes_with_zips <- read.csv("Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset

# Secondary dataset - Median Income by ZCTA
cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")

@@

# Secondary dataset - COVID cases in Los Angeles County
covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")

evictions_la <- read.csv("Raw_Data/LA_evictions.csv")
```

@@

```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week"}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

## Median Household Income by Zip Code in Los Angeles by Year

```{r, echo=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

@@

# taking a random sample to reduce load on model
#set.seed(123)
#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
#rownames(sampled_ids) <- NULL
#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
#write.csv(zip_codes, "sampled_crimes.csv")
zip_codes <- read.csv("sampled_crimes.csv")

@@

@@

## Number of Evictions and Crime Levels Per Month
```{r, echo=FALSE}
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

@@

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```
### Interpretation
This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

## Boxplot - Kaitlyn

```{r, echo=FALSE}
crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()

bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")

@@

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau))
```

## geom_tile for correlations - Bianca

@@

# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
LA_county_evictions <- read.csv("Raw_Data/LA_county_evictions.csv")

@@

# next, unemployment rate
unemployment <- read.csv("Raw_Data/la_county_unemployment_rate.csv")
unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))

# lapd budget
lapd_budget <- read.csv("Raw_Data/lapd_budget.csv")

@@

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) + 
  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
```
## geom_jitter

```{r, echo=FALSE}
ggplot(crime_la, aes(x = DateParsed, y = Vict.Sex, color = Vict.Sex)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(title = "Crime Over Time by Gender",
       x = "Date",
       y = "Gender") +
  theme_minimal()

## Density Plot

crime_la$Hour <- as.numeric(format(crime_la$DateParsed, "%H"))

ggplot(crime_la, aes(x = Hour, fill = ..density..)) +
  geom_density(alpha = 0.5, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Crime Density Per Hour",
       x = "Hour",
       y = "Density") +
  theme_minimal()

## Plot 6

@@

## Data Wrangling {.unnumbered}

# Reading in Raw Datasets

@@

# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")

@@

```{r Combining Median Income, eval=FALSE}
# Secondary datasets
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

@@

### Zip Codes for Crimes {.unnumbered}
```{r Get Zip Codes, eval=FALSE}
library(sf)
library(dplyr)

crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")

@@

zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")

@@

  write.csv(result, "crime_zip_files/all_zips.csv")

crime_zctas <- read.csv("Raw_Data/all_zips.csv")
colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")

@@

# Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}

@@

# Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}

@@

# Figure 3 - Crime rate frequency by day of the week {.unnumbered}

@@

# Figure 4 - Gender Distribution for crimes over time {.unnumbered}

@@

# Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}

```{r Figure 5 Wrangling, eval=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

# Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}

@@

# taking a random sample to reduce load on model
#set.seed(123)
#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
#rownames(sampled_ids) <- NULL
#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
#write.csv(zip_codes, "sampled_crimes.csv")
zip_codes <- read.csv("sampled_crimes.csv")

@@

## Crime \~ Week Regression Code {.unnumbered}

@@

## Crime \~ Income Regression Code {.unnumbered}

>>>>

author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
fontsize: 10pt
toc: true                   # add table of contents at the beginning
toc-depth: 3               
geometry: margin=1in
bibliography: references.bib
format:
  pdf: default
number-sections: true
---

@@

# primary datasets
# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")

crime_la <- read.csv("0_Data/Cleaned_Data/all_crimes_2010_present.csv")

# primary dataset zip code and geometry data
crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset

# Secondary dataset - Median Income by ZCTA
cali_median_income <- read.csv("0_Data/Cleaned_Data/cali_median_income.csv")

@@

# Secondary dataset - COVID cases in Los Angeles County
covid_14day <- read.csv("0_Data/Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")

evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
```

@@

```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week", message=FALSE}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

## Median Household Income by Year and Zip Code in Los Angeles

```{r, echo=FALSE}
zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

# color palette for sequential data
palette <- brewer.pal(7, "Reds")

@@

# taking a random sample to reduce load on model
#set.seed(123)
#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
#rownames(sampled_ids) <- NULL
#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
#write.csv(zip_codes, "sampled_crimes.csv")
zip_codes <- read.csv("0_Data/Cleaned_Data/sampled_crimes.csv")

@@

plot(x = crime_income_per_zip$cali_median_income, 
     y = crime_income_per_zip$number_of_crimes, 
     type = "p",
     ylab = "Number of Crimes",  
     xlab = "Median Income ($)",  
     bty = "l")
```

@@

## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025
```{r, echo=FALSE}
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

@@

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```
### Interpretation
This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

## Annual Number of Crimes by Los Angeles Police Department Bureau

```{r, echo=FALSE}
crime_la$year <- lubridate::year(crime_la$DateParsed)
crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()

bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")

@@

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau)) +
  labs(y = "Annual Number of Crimes")
```
### Interpretation

This boxplot shows the number of crimes in the four different bureaus of Los Angeles annually. We categorized the areas into their respective Central, South, Valley, and West Bureaus. The South Bureau has the highest median number of crimes, as well as the largest variability, while the Central, Valley, and West bureaus have relatively similar distributions with lower crime counts.The outliers suggests some years had significantly lower crime counts than others. This visualization helps us understand how crime is distributed across different regions, with the South experiencing the highest fluctuation and overall crime levels.

## Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors

@@

# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")

@@

# next, unemployment rate
unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))

# lapd budget
lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")

@@

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```
### Interpretation
This heat map shows the correlations between different variables that could affect the number of crimes committed. Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles. The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.

Here, we see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions. Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased. Similarly, as the number of evictions increases, so does the number of crimes. Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime). The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.

## Homicide Victims by Age and Gender in Los Angeles 2010-Present

```{r, echo=FALSE}
homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))

levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(x = "Gender",
       y = "Age") +
  theme_minimal() +
  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
  theme(legend.position = 'none')
```
### Interpretation

This jitter plot shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present. Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65. Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed. This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.

## Frequency of Crimes in Los Angeles Throughout the Day

```{r, echo=FALSE}
crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))

ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(x = "Hour (Military Time)",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = 'none')
```
### Interpretation

This density plot shows the frequency of crimes depending on the hour (in military time) that they occurred. Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM). The data peaks at noon, meaning the most crimes occur in that hour of the day. Crimes are more frequent after that, until the early morning. The second most common crime occurrence is 6 PM, or 18 hours.

## Time series analyzing homicide rate trends over time by socioeconomic quartiles

```{r, echo=FALSE}
# Data Preparation
homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
  mutate(date = as.Date(DateParsed),
         zip = as.character(ZCTA5CE20)) %>% select(-X)

homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
          ref = mean(S1903_C03_001E),
          .groups = "drop")

homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
# removing NAs
homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile",
    title = "Homicide Rates in Los Angeles By Median Income Level"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)
```

### Interpretation
Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic. In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic. On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.

@@

#### Data Wrangling {.unnumbered}

##### Reading in Raw Datasets

@@

# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")

@@

```{r Combining Median Income, eval=FALSE}
# Secondary datasets
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("0_Data/Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

@@

##### Zip Codes for Crimes
```{r Get Zip Codes, eval=FALSE}
library(sf)
library(dplyr)

crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")

crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")

@@

zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")

@@

  write.csv(result, "0_Data/Raw_Data/all_zips.csv")

crime_zctas <- read.csv("0_Data/Raw_Data/all_zips.csv")
colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")

@@

#### Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}

@@

#### Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}

@@

#### Figure 3 - Crime rate frequency by day of the week {.unnumbered}

@@

#### Figure 4 - Gender Distribution for crimes over time {.unnumbered}

@@

#### Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}

```{r Figure 5 Wrangling, eval=FALSE}
zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

#### Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}

@@

# taking a random sample to reduce load on model
#set.seed(123)
#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
#rownames(sampled_ids) <- NULL
#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
#write.csv(zip_codes, "sampled_crimes.csv")
zip_codes <- read.csv("0_Data/sampled_crimes.csv")

@@

####Figure 7 - Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025 {.unnumbered}

```{r, eval=FALSE}
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```

#### Figure 8 - Annual Number of Crimes by Los Angeles Police Department Bureau {.unnumbered}

```{r, eval=FALSE}
crime_la$year <- lubridate::year(crime_la$DateParsed)
crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()

bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau)) +
  labs(y = "Annual Number of Crimes")
```

#### Figure 9 - Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors {.unnumbered}

```{r, eval=FALSE}
library(lubridate)
# creating dataset for correlations

# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")

# next, unemployment rate
unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))

# lapd budget
lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```

#### Figure 10 - Homicide Victims by Age and Gender in Los Angeles 2010-Present {.unnumbered}

```{r, eval=FALSE}
homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))

levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(x = "Gender",
       y = "Age") +
  theme_minimal() +
  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
  theme(legend.position = 'none')
```

#### Figure 11 - Frequency of Crimes in Los Angeles Throughout the Day {.unnumbered}

```{r, eval=FALSE}
crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))

ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(x = "Hour (Military Time)",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = 'none')
```

#### Figure 12 - Time series analyzing homicide rate trends over time by socioeconomic quartiles {.unnumbered}

```{r, eval=FALSE}
# Data Preparation
homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
  mutate(date = as.Date(DateParsed),
         zip = as.character(ZCTA5CE20)) %>% select(-X)

homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
          ref = mean(S1903_C03_001E),
          .groups = "drop")

homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
# removing NAs
homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile",
    title = "Homicide Rates in Los Angeles By Median Income Level"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)
```
#### Crime \~ Week Regression Code {.unnumbered}

@@

#### Crime \~ Income Regression Code {.unnumbered}
